---
title: "P8110 - Applied Regression II - Homework #3"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
library(survival)
library(survminer)
library(tidyverse)
```

The ”HW3Data” gives the time until staphylococcus infection (in days) for 154 patients with a burn wound. The purpose of this study was to compare a routine bathing care treatment with a body-cleansing method. The time until staphylococcus infection was recorded, along with an indicator variable – whether or not an infection had occurred. The severity of the burn was measured by percentage of total surface area of body burned.  

|variable             |description                                              |
|---------------------|---------------------------------------------------------|
|id                   |patient id                                               |
|treatment            |0 = routine bathing <br> 1 = body cleansing              |
|pct_burned           |percentage of total surface area burned                  |
|time_to_infection    |time to staphylocous aureaus infection or on study time  |
|infection            |0 = infection did not occur <br> 1 = infection occurred  |  

***  

### Question 1  

Use R to compute the Kaplan-Meier estimator and 95% CI of the survival function for the routine bathing group, and also for the body-cleansing group (only keep the K-M estimates for unique event time points). Generate a graph of the survival functions in the two treatment groups. Interpret the graph. (*Hint: What do you observe about the difference of the survival functions between the two groups?*)  

```{r, message = FALSE}
hw3_data <- read_csv("./data/hw3_data.csv",
                     col_names = c("id", "treatment", "pct_burned", "time_to_infection", "infection"))

hw3_data
```

#### Estimate the survival curve  

```{r}
q1_fit <- survfit(Surv(time = time_to_infection, event = infection) ~ treatment, data = hw3_data)
```

#### Kaplan-Meier Estimators

##### Routine Bathing Group  

```{r}
q1_fit %>% 
    broom::tidy() %>% 
    filter(strata == "treatment=0" & !n.event == 0) %>% 
    knitr::kable(digits = 5) %>% 
    print(n = Inf)
```

##### Body-Cleansing Group  

```{r}
q1_fit %>% 
    broom::tidy() %>% 
    filter(strata == "treatment=1" & !n.event == 0) %>% 
    knitr::kable(digits = 5) %>% 
    print(n = Inf)
```

##### Survival Curve Plot  

```{r}
ggsurvplot(q1_fit, conf.int = TRUE)
```

***  

### Question 2  

Test whether the survival curves for the two treatment groups are the same using both the log-rank test and Wilcoxon test at *alpha* = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion for each test. Explain the difference between the two tests.  

##### Log-Rank Test  

```{r}
survdiff(Surv(time = time_to_infection, event = infection) ~ treatment, data = hw3_data) %>% 
    broom::glance() %>% 
    knitr::kable(
        col.names = c("test statistic", "degrees of freedom", "p-value"),
        align = "c",
        digits = 3)
```

##### Wilcoxon Test  

```{r}
survdiff(Surv(time = time_to_infection, event = infection) ~ treatment, data = hw3_data, rho = 1) %>% 
    broom::glance() %>% 
    knitr::kable(
        col.names = c("test statistic", "degrees of freedom", "p-value"),
        align = "c",
        digits = 3)
```

### Question 3  

Test whether the survival functions are the same among the four groups using the generalized log-rank test at *alpha* = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion.  


